---
title: "project"
author: "zhu,zhaoyang"
date: "2019.1.22"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
if (! require(readr)){
  install.packages("readr")
}
if (! require(tidyverse)){
  install.packages("tidyverse")
}
if (! require(coin)){
  install.packages("coin")
}
if (! require(dplyr)){
  install.packages("dplyr")
}
if (! require(rcompanion)){
  install.packages("rcompanion")
}
if (! require(MASS)){
  install.packages("MASS")
  }
if (! require(AER)){
  install.packages("AER")
}
knitr::opts_chunk$set(echo = TRUE)
data = read_csv("Modified1.csv", 
                col_types = cols(`Channel` = col_skip()))
```

# First look at the data.

```{r,echo=TRUE,eval=TRUE}
head(data)
group = as.factor(data$Group)
data2 = data
data2$Group = group
data2$Duration = as.numeric(data2$Duration)
data2$Duration[data2$Duration == 0] = NA
data2$Subject = as.factor(data2$Subject)
data2 = filter(data2, !is.na(Duration))
data2 = data2[data2$Duration >= 5, ]
summary(data2)
par(mfrow = c(1,2))
summary(data2$Subject)
```

# EDA of the data after cleaning.

```{r,echo=TRUE,eval=TRUE}
subject = c(9,10,16,35,36,43,45,55,68,69,70,83)

# ave duration
aveDuration = c()
for (i in subject){
  this = mean(data2$Duration[data2$Subject == i], na.rm = TRUE)
  if (is.na(this)){
    this = 0
  }
  aveDuration = c(aveDuration, this)
  
}
Frequency = c()
for (i in subject){
  count = length(data2$Subject[data2$Subject == i])
  if (count == 1){
    count = 0
  }
  Frequency = c(Frequency, count)
  
}
RR = c()
for ( i in subject){
  RR = c(RR, data$RR[data$Subject == i][1])
}

aveDurationData = tibble("aveDuration" = aveDuration, "subject" = subject, "frequency" = Frequency, "RR" = RR, "group" = c("S","S","S","T","T","M","M","M","L","L","L","T"))
aveDurationData$group = as.factor(aveDurationData$group)
aveDurationData$subject = factor(aveDurationData$subject, c(9,10,16,35,36,83,43,45,55,68,69,70), order = TRUE)
ggplot(data=aveDurationData) + geom_point(aes(x = group, y = aveDuration)) + ylab("average seizure duration")
ggplot(data=aveDurationData) + geom_point(aes(x= subject, y = frequency))
ggplot(data=aveDurationData) + geom_point(aes(x = group, y = frequency))
ggplot(data=aveDurationData) + geom_point(aes(x = subject, y = RR)) + ylab("Righting Reflex")
ggplot(data=aveDurationData) + geom_boxplot(aes(x = subject, y = aveDuration)) + ylab("average seizure duration")
```

From the plots we can see that potential ouliers subjec 10 and subject 45

at the same time, there is difference in groups.


# linear regression models

```{r, echo=TRUE, eval = TRUE}
model1 = lm(aveDurationData$frequency ~ aveDurationData$group + aveDurationData$RR)
model2 = lm(aveDurationData$aveDuration ~ aveDurationData$group + aveDurationData$RR)
plot(model2)
plot(model1)
```

First try linear regression on dependent variables.

From the residual and normal QQ plots, we can see that our assumptions are violated.

# Non-parametric tests
```{r, echo=TRUE, eval=TRUE}
# non-parametric tests
median_test(data = aveDurationData, aveDuration ~ group) # No evidence showing that the group medians are different
median_test(data = aveDurationData, frequency ~ group) # No evidence showing that the group medians are different
pairwisePermutationTest(data = aveDurationData, aveDuration ~group) # No evidence showing that the group are different
pairwisePermutationTest(data = aveDurationData, frequency ~group) # No evidence showing that the group are different
kruskal.test(data=aveDurationData, frequency ~ group) # No evidence showing that the group are different
kruskal.test(data=aveDurationData, aveDuration ~ group) # No evidence showing that the group are different
```

Since our data is not normal distributed, we use some non-parametric tests:

Moon’s median test is a median based test, the influence of outliers is limited in this test. 

Pairwise permutation test resamples the observed data many times to determine a p-value for the test, it is an extension of permutation test to multiple groups.  

Kruskal–Wallis test is a rank-based test.

Assumptions are :

1. Dependent variable is ordinal, interval, or ratio.

2. Independent variable is a factor with levels indicating groups.

3. Observations between groups are independent.  

And we satisfied them all.

```{r, echo=TRUE, eval=TRUE}
# poisson regression on frequency
model3 = glm(aveDurationData$frequency ~ aveDurationData$group + aveDurationData$RR, family = "poisson")
summary(model3)
aveDurationData$group = relevel(aveDurationData$group, ref = "S")
model4 = glm(aveDurationData$frequency ~ aveDurationData$group + aveDurationData$RR, family = "poisson")
summary(model4)
aveDurationData$group = relevel(aveDurationData$group, ref = "M")
model5 = glm(aveDurationData$frequency ~ aveDurationData$group + aveDurationData$RR, family = "poisson")
summary(model5)

# poisson regression with outliers but no righting reflex
model6 = glm(aveDurationData$frequency ~ aveDurationData$group, family = "poisson")
summary(model6)
aveDurationData$group = relevel(aveDurationData$group, ref = "S")
model7 = glm(aveDurationData$frequency ~ aveDurationData$group, family = "poisson")
summary(model7)
aveDurationData$group = relevel(aveDurationData$group, ref = "L")
model8 = glm(aveDurationData$frequency ~ aveDurationData$group, family = "poisson")
summary(model8)

# poisson regression with subject 45 and right reflex but remove subject 10
no_ten = filter(aveDurationData, frequency != 43)
model9 = glm(data = no_ten, frequency ~ group + RR, family = "poisson")
summary(model9)
no_ten$group = relevel(no_ten$group, ref = "M")
model10 = glm(data = no_ten, frequency ~ group + RR, family = "poisson")
summary(model10)
no_ten$group = relevel(no_ten$group, ref = "S")
model11 = glm(data = no_ten, frequency ~ group + RR, family = "poisson")
summary(model11)

# poisson regression with subject 45 but not RR or subject 10
model9 = glm(data = no_ten, frequency ~ group, family = "poisson")
summary(model9)
no_ten$group = relevel(no_ten$group, ref = "M")
model10 = glm(data = no_ten, frequency ~ group, family = "poisson")
summary(model10)
no_ten$group = relevel(no_ten$group, ref = "L")
model11 = glm(data = no_ten, frequency ~ group, family = "poisson")
summary(model11)

# take out the outlier subject 45
no_outlier_data = filter(aveDurationData, frequency != max(frequency))
# poisson regression with no outlier 45
poisson_no_out_one = glm(data = no_outlier_data, frequency ~ group + RR, family = "poisson")
summary(poisson_no_out_one)
no_outlier_data$group = relevel(no_outlier_data$group, ref = "M")
poisson_no_out_two = glm(data = no_outlier_data, frequency ~ group + RR, family = "poisson")
summary(poisson_no_out_two)
no_outlier_data$group = relevel(no_outlier_data$group, ref = "S")
poisson_no_out_three = glm(data = no_outlier_data, frequency ~ group + RR, family = "poisson")
summary(poisson_no_out_three)

# No outlier 45 and 10 but with Righting Reflex
no_outlier_data_two = filter(no_outlier_data, frequency != 43)
poisson_no_out_four = glm(data = no_outlier_data_two, frequency ~ group + RR, family = "poisson")
summary(poisson_no_out_four)
no_outlier_data_two$group = relevel(no_outlier_data_two$group, ref = "M")
poisson_no_out_five = glm(data = no_outlier_data_two, frequency ~ group + RR, family = "poisson")
summary(poisson_no_out_five)
no_outlier_data_two$group = relevel(no_outlier_data_two$group, ref = "L")
poisson_no_out_six = glm(data = no_outlier_data_two, frequency ~ group + RR, family = "poisson")
summary(poisson_no_out_six)

# No righting reflex and subject 45 but with subject 10
poisson_no_out_seven = glm(data = no_outlier_data, frequency ~ group, family = "poisson")
summary(poisson_no_out_seven)
no_outlier_data$group = relevel(no_outlier_data$group, ref = "M")
poisson_no_out_eight = glm(data = no_outlier_data, frequency ~ group, family = "poisson")
summary(poisson_no_out_eight)
no_outlier_data$group = relevel(no_outlier_data$group, ref = "S")
poisson_no_out_nine = glm(data = no_outlier_data, frequency ~ group, family = "poisson")
summary(poisson_no_out_nine)
```

Since our dependent variable frequency is a count, we can fit a poisson regression.

we split into cases to examine the impacts of outliers and righting reflex on our Poisson regression model.

We change reference group to compare different groups.


# Linear regression
```{r, echo=TRUE, eval=TRUE}
# linear regression with no outlier
# still no difference between groups
linear_no_out = lm(data = no_outlier_data_two, aveDuration ~ group + RR)
plot(linear_no_out)
```

Fit linear regression model again to see if assumptions are satisifed without outliers.

However, assumptions are still violated


# Negative Binomial Regression
```{r, echo=TRUE, eval=TRUE}
# evidence of over-dispersion
AER::dispersiontest(model3)
# over disperson and under disperson
nb_model = glm.nb(data = aveDurationData, formula = frequency ~ group + RR)
summary(nb_model)
aveDurationData$group = relevel(aveDurationData$group, ref = "S")
nb_model2 = glm.nb(data = aveDurationData, formula = frequency ~ group + RR)
summary(nb_model2)
aveDurationData$group = relevel(aveDurationData$group, ref = "M")
nb_model3 = glm.nb(data = aveDurationData, formula = frequency ~ group + RR)
summary(nb_model3)


```

over-dispersion observed in our Poisson model, so we decided to fit a negative binomial regression.
